---
title: "Framingham Heart Study Visualization"
author: "Al Shain"
date: "December 25, 2015"
output: html_document
---
```{r message = FALSE}
library(dplyr)
library(ggplot2)
library(reshape2)
library(psych)
setwd("~/Desktop/Coursera/DevelopingDataProducts/CourseProject-ShinyAppReproduciblePitch")
```

## The Framingham Heart Study (FHS)
* A long term study of Cardiovascular Disease among subjects in Framingham, USA.
* It began in 1948 with 5209 subjects.
* Patients have been examined every 2 years for
    * Clinical examination data  
    * Cardiovascular outcomes  
* More information about the study can be found at the FHS
  [official website](http://www.framinghamheartstudy.org/)
  
## Some Findings of the FHS
* Over 1000 medical papers have been published using the data from the FHS.
* The study has uncovered or confirmed several risk factors for Heart Disease:
    * High blood pressure
    * Obesity
    * Cigarette smoking
* The study has also shown there are differences in coronary risk between
  men and women.
    
## The Framingham Dataset
> - We have a subset of the original dataset, pertaining to 4434 patients.
> - Each of these patients was followed for a total of 24 years, but we only have
  data on their clinical examinations for a period of 12 years, spanning 1956 to 1968. 
    * The data is provided in longitudinal form.
    * Each participant has 1 to 3 observations.
    * There are a total of 11627 rows in the data.

## Columns in the Framingham Dataset
Here are the *first six columns and rows* of the dataset:
```{r echo=FALSE, label="load data"}
fhs.long <- read.csv("/Users/Al/Desktop/Coursera/DevelopingDataProducts/CourseProject-ShinyAppReproduciblePitch/data/frmgham2.csv")
head(fhs.long[,1:6])
```
> - `RANDID` refers to the patient id.
> - The remaining columns are measurements on the patients during their
  examinations, and outcome measures. 

## Measurements in Framingham Dataset
Variable | Explanation
---------|------------
SEX    | 1 for males and 2 for females.   
PERIOD | An integer between 1 and 3 denoting the examination number.  
AGE    | Age at examination (in years).  
SYSBP, DIABP |  Systolic and diastolic blood pressure.   

## Measurements in Our Dataset | cont'd
Variable | Explanation
---------|------------
CIGPDAY | Number of cigarettes smoked per day.   
BMI | Body Mass Index.
LDLC | Low density cholesterol.
TOTCHOL | Total cholesterol.
HYPERTEN | 1 if patient has hypertension, which is defined to be Systolic greater than 140mmHg or Diastolic greater than 90mmHg.

## Outcome Measures in Our Dataset
Variable | Explanation
---------|------------
STROKE | 1 if patient experienced a stroke during the 12 years and 0 otherwise.   
ANYCHD | 1 if patient experienced a Coronary Heart Disease during the 12 years and 0 otherwise.   
CVD    | 1 if patient experienced a Cerebrovascular Disease during the 12 years and 0 otherwise. 

# Framingham Graphs
#### We wish to know whether the mean heart rate of "healthy" people at exam 1 is equal to 70 beats per minute (bpm). 

```{r, label="ex 4 soln"}
fhs <- read.csv("/Users/Al/Desktop/Coursera/DevelopingDataProducts/CourseProject-ShinyAppReproduciblePitch/data/frmgham2.csv")
colnames(fhs) <- toupper(colnames(fhs))
with(fhs, summary(HEARTRTE))
ggplot(fhs, aes(HEARTRTE)) +
  geom_histogram(color = "black", fill = "white")
#
fhs.healthy <- filter(fhs, PERIOD == 1 & CURSMOKE == 0 & BMI < 25 & DIABP < 80 
                      & SYSBP < 120 & DIABETES == 0)
with(fhs.healthy, describe(HEARTRTE))
ggplot(fhs.healthy, aes(x = HEARTRTE, y = ..density..)) + 
  geom_histogram(color = "black", fill = "white") +
  geom_density(fill = "green", alpha = 0.2)
fhs.wide <- dcast(fhs, RANDID + SEX ~ PERIOD, value.var = "HEARTRTE")
colnames(fhs.wide)[3:5] <- c("HEARTRTE1", "HEARTRTE2", "HEARTRTE3")
tbl_df(fhs.wide)
set.seed(12345)
fhs.asample <- sample_n(fhs.wide, 200)
ggplot(fhs.asample, aes(x = HEARTRTE1 - HEARTRTE2, y = ..density..)) + 
  geom_histogram(color = "black", fill = "white") +
  geom_density(fill = "red", alpha = 0.2)
ggplot(fhs.wide, aes(HEARTRTE1, fill = as.factor(SEX))) + 
  geom_density(alpha = 0.2)
ggplot(filter(fhs, PERIOD == 1), aes(x = as.factor(EDUC), y = BMI)) + 
  geom_boxplot() +
  stat_summary(fun.y = "mean", geom = "point", shape = 4)
```

# Shiny Product
## Input Panel 1
```{r eval=FALSE, label="input panel 1"}
inputPanel(
  radioButtons("x_var", label="Explanatory Variable", 
    choices=c("BMI", "LDLC", "TOTCHOL"), selected="BMI"),
  
  radioButtons("y_var", label="Response Variable", 
    choices=c("CVD", "STROKE", "ANYCHD"), selected="CVD"),
  
  checkboxInput("con_gender", label="Condition on Gender", value=FALSE)
)
```

# Shiny Product
## Input Panel 2
```{r eval=FALSE, label="input panel 2"}
inputPanel(
  sliderInput("bmi_range", label="BMI range", min=10,
  max=60, value=c(10,60)),
  sliderInput("ldlc_range", label="LDLC range", min=20,
  max=565, value=c(20,565)),
  sliderInput("tot_range", label="Total Chol. range", min=112,
  max=625, value=c(112,625))
)
```
 
## Output Panel
```{r eval=FALSE, label="output panel 1"}
  x <- fhs.only3[,input$x_var]
  y <- fhs.only3[,input$y_var]
  
  if(input$x_var == "BMI") {
    x.range <- input$bmi_range
  } else if (input$x_var == "LDLC") {
    x.range <- input$ldlc_range
  } else 
    x.range <- input$tot_range
```

## Output Panel
```{r eval=FALSE, label="output panel 2"}
  if(input$con_gender){
    p <- qplot(x,y, xlab=input$x_var, ylab=input$y_var, facets = . ~ SEX, data=fhs.only3)
  } else {
    p <- qplot(x,y, xlab=input$x_var,ylab=input$y_var) 
  } 
  p + xlim(x.range) + stat_smooth(method="glm", family="binomial") + 
  geom_point(position=position_jitter(height=0.02)) 
```




